---
title: "Reproducible Research Project 1"
author: "Priyanka Roy"
date: "February 10, 2017"
output: html_document
---
Reading the required packages
```{r}
library(plyr)
library(ggplot2)
library(lattice)
```

Reading the data:
```{r, echo= TRUE}
activity <- read.csv("D:/Priyanka/Coursera/5.ReproducibleResearch/activity.csv")
head(activity)
```

Calculating total number of steps taken per day, its mean and median as well
```{r}
activity_sum <- ddply(activity, "date",summarise, StepsTaken = sum(steps))
activity_sum <- na.omit(activity_sum)
hist(activity_sum$StepsTaken, main = "Steps taken per day", col = "red", xlab = "Number of steps")
mean(activity_sum$StepsTaken)
median(activity_sum$StepsTaken)
```

Average daily activity pattern
```{r}
interval_max <- ddply(activity, "interval", summarise, Max_Interval = mean(steps, na.rm = TRUE))
plot(interval_max$interval, interval_max$Max_Interval, type = "l", main = "Average Steps on interval")
interval_max$interval[interval_max$Max_Interval ==  max(interval_max$Max_Interval)]
```

Imputing Missing values:
Total number of missing vlues is:
```{r}
sum(is.na(activity))
```
Strategy- I am replacing all the null values by the average number of steps taken in the corresponding 5-minute interval
as calculated in interval_max table. New dataset is below:
```{r}
activity_imputed <- activity
for(i in 1:nrow(activity_imputed)){
  activity_imputed$steps[i]<- ifelse(is.na(activity_imputed$steps[i]),
                              interval_max$Max_Interval[interval_max$interval== activity_imputed$interval[i]],
                              activity_imputed$steps[i])
}
activity_imputed$steps <- round(activity_imputed$steps,0)
```

plot and comparison of new data with old data
```{r}
act_imp_sum <-  ddply(activity_imputed, "date",summarise, StepsTaken = sum(steps))
hist(act_imp_sum$StepsTaken, main = "Steps taken per day by Imputed data", col = "red", xlab = "Number of steps")
mean(act_imp_sum$StepsTaken)
median(act_imp_sum$StepsTaken)
```
After imputing the data, mean is almost the same. However, median got reduced by 4 units.

Differences in activity patterns between weekdays and weekends
Week Factor created:
```{r}
activity_imputed$NewDate <- weekdays(as.Date(activity_imputed$date))
activity_imputed$Week <- ifelse(activity_imputed$NewDate %in% c("Sunday","Saturday"),"weekend", "weekdays")
```
plot for weekend/weekday
```{r}
activity_imputed_agg <- ddply(activity_imputed, c("Week","interval"),summarise, Steps = mean(steps, na.rm = TRUE))
xyplot(Steps ~ interval | Week,data = activity_imputed_agg, type = "l", xlab = "Interval", ylab = "Steps Taken",layout = c(1,2), main = "Average Steps taken in 5-minute interval(Weekends/Weekdays)")
```



